---
title: "Journal (reproducible report)"
author: "Ahmed Elkhateeb"
date: "2020-11-27"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

**IMPORTANT:** You can delete everything in here and start fresh. You might want to start by not deleting anything above this line until you know what that stuff is doing.

This is an `.Rmd` file. It is plain text with special features. Any time you write just like this, it will be compiled to normal text in the website. If you put a \# in front of your text, it will create a top level-header.

# My first post

Last compiled: `r Sys.Date()`

Notice that whatever you define as a top level header, automatically gets put into the table of contents bar on the left. 

## Second level header

You can add more headers by adding more hashtags. These won't be put into the table of contents

### third level header

Here's an even lower level header

# My second post (note the order)

Last compiled: `r Sys.Date()`

I'm writing this tutorial going from the top down. And, this is how it will be printed. So, notice the second post is second in the list. If you want your most recent post to be at the top, then make a new post starting at the top. If you want the oldest first, do, then keep adding to the bottom

# Adding R stuff

So far this is just a blog where you can write in plain text and serve your writing to a webpage. One of the main purposes of this lab journal is to record your progress learning R. The reason I am asking you to use this process is because you can both make a website, and a lab journal, and learn R all in R-studio. This makes everything really convenient and in the same place. 

So, let's say you are learning how to make a histogram in R. For example, maybe you want to sample 100 numbers from a normal distribution with mean = 0, and standard deviation = 1, and then you want to plot a histogram. You can do this right here by using an r code block, like this:

```{r}
# Data Science at TUHH ------------------------------------------------------

# Chapter 2 ----
# SALES ANALYSIS ----

# 2.1 Load libraries ----
  library(tidyverse)
  library(lubridate)
  library(readxl)

# 2.2 Importing Files ----
  bikes_tbl      <- read_excel("data-science/00_data/01_bike_sales/01_raw_data/bikes.xlsx")
  orderlines_tbl <- read_excel("data-science/00_data/01_bike_sales/01_raw_data/orderlines.xlsx")
  bikeshops_tbl  <- read_excel("data-science/00_data/01_bike_sales/01_raw_data/bikeshops.xlsx")

# 2.3 Joining Data ----
  bike_orderlines_joined_tbl <- orderlines_tbl %>%
  left_join(bikes_tbl, by = c("product.id" = "bike.id")) %>%
  left_join(bikeshops_tbl, by = c("customer.id" = "bikeshop.id"))

# 2.4 Wrangling Data ----
  bike_orderlines_wrangled_tbl <- bike_orderlines_joined_tbl %>%
    # 2.4.1 Split location into state and city
    separate(col    = location,
            into   = c("city", "state"),
            sep    = ", ") %>%

    # 2.4.2 Add the total price (price * quantity) 
    mutate(total.price = price * quantity) %>%
  
    # 2.4.3 Rename columns in underscores instead of dots
    set_names(names(.) %>% str_replace_all("\\.", "_"))

# 2.5 Business Insights ----
  
  # 2.5.1 Sales by State
  # Step 1 - Manipulate
  sales_by_location_tbl <- bike_orderlines_wrangled_tbl %>%
  
    # Select columns
    select(state, total_price) %>%
  
    # Grouping by state and summarizing sales
    group_by(state) %>% 
    summarize(sales = sum(total_price)) %>%
  
    # Turn the numbers into a currency format 
    # mutate(sales_text = scales::dollar(sales)) <- Works for dollar values
    mutate(sales_text = scales::dollar(sales, big.mark = ".", 
                                     decimal.mark = ",", 
                                     prefix = "", 
                                     suffix = " €"))

  # Step 2 - Visualize
  sales_by_location_tbl %>%
  
    # Setup canvas with the columns year (x-axis) and sales (y-axis)
    ggplot(aes(x = state, y = sales)) +
  
    # Geometries
    geom_col(fill = "#2DC6D6") + # Use geom_col for a bar plot
    geom_label(aes(label = sales_text)) + # Adding labels to the bars
    geom_smooth(method = "lm", se = FALSE) + # Adding a trendline
  
    # Formatting
    # scale_y_continuous(labels = scales::dollar) + # Change the y-axis. 
    # Again, we have to adjust it for euro values
    scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
    labs(
      title    = "Revenue by state",
      subtitle = "",
      x = "", # Override defaults for x and y
      y = "Revenue"
    ) +
  
    #rotate x-axis labels
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  # 2.5.2 Sales by Year and Location

  # Step 1 - Manipulate
  sales_by_year_location_tbl <- bike_orderlines_wrangled_tbl %>%
  
    # Select columns and add a year
    select(order_date, total_price, state) %>%
    mutate(year = year(order_date)) %>%
  
    # Group by and summarize year and main catgegory
    group_by(year, state) %>%
    summarise(sales = sum(total_price)) %>%
    ungroup() %>%
  
    # Format $ Text
    mutate(sales_text = scales::dollar(sales, big.mark = ".", 
                                      decimal.mark = ",", 
                                      prefix = "", 
                                      suffix = " €"))

  # Step 2 - Visualize
  sales_by_year_location_tbl %>%
  
    # Set up x, y, fill
    ggplot(aes(x = year, y = sales, fill = state)) +
  
    # Geometries
    geom_col() + # Run up to here to get a stacked bar plot
  
    # Facet
    facet_wrap(~ state) +
  
    # Formatting
    scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
    labs(
      title = "Revenue by year and location",
      subtitle = "",
      fill = "States" # Changes the legend name
    ) +

    #rotate x-axis labels
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  # Chapter 3 ----

# 3.1 API data retrieval ----

# 3.1.0 Load libraries

library(httr)
library(glue)
library(jsonlite)
library(keyring)

# 3.1.1 Code
#Returns information for the top 50 cities worldwide.

#keyring::key_set("token")

api_key = key_get("token")

url <- modify_url(url = "http://dataservice.accuweather.com", path = glue("/locations/v1/topcities/50?apikey={api_key}&language=en-us&details=false"))

raw_data <- GET(url)

stop_for_status(raw_data) # automatically throws an error if a request did not succeed

data_list <- raw_data %>% 
  .$content %>% 
  rawToChar() %>% 
  fromJSON()

glimpse(data_list)

# 3.2 WEBSCRAPING ----

# 3.2.0 Load libraries

library(tidyverse) # Main Package - Loads dplyr, purrr, etc.
library(rvest)     # HTML Hacking & Web Scraping
library(xopen)     # Quickly opening URLs
library(jsonlite)  # converts JSON files to R objects
library(glue)      # concatenate strings
library(stringi)   # character string/text processing

# 3.2.1 COLLECT PRODUCT FAMILIES

url_home          <- "https://www.rosebikes.de/"

# Read in the HTML for the entire home webpage
html_home         <- read_html(url_home)

# Web scrape the urls for the families
bike_family_tbl <- html_home %>%
  
  # Get the nodes for the families ...
  html_nodes(css = ".main-navigation-category-with-tiles__link ") %>%
  
  # ...and extract the href information
  html_attr('href') %>%
  
  # Remove the product families sale 
  discard(.p = ~stringr::str_detect(.x,"sale$")) %>%
  
  # Convert vector to tibble
  enframe(name = "position", value = "family_class") %>%
  
  # Save the family urls
  transmute(family_url = str_glue("https://www.rosebikes.de{family_class}"))


# 3.2.2 COLLECT Gravel Categories 

# Read in the HTML for the entire Gravel family webpage
html_family         <- read_html(bike_family_tbl$family_url[3])


bike_category_tbl <- html_family %>%
  
  #Get the nodes for categories
  html_nodes(css = ".catalog-category-bikes__picture-wrapper") %>%
  
  # ...and extract the href information
  html_attr('href') %>%
  
  # Convert vector to tibble
  enframe(name = "position", value = "subdirectory") %>%
  
  # Save the Categories urls
  transmute(position, category_url = str_glue("https://www.rosebikes.de{subdirectory}"))


# 3.2.3 COLLECT Gravel/Backroad Bike Models 

# Read in the HTML for the entire Gravel/Backroad category webpage
html_models         <- read_html(bike_category_tbl$category_url[2])

bike_model_tbl <- html_models %>%
  
  #Get the nodes for bike models
  html_nodes(css = ".catalog-category-model__title") %>%
  
  #...and extract the text
  html_text()%>%
  
  # Convert vector to tibble
  enframe(name = "position", value = "model")


#Read models prices
model_price_tbl <- html_models %>%
  
  #Get the nodes for bike models prices
  html_nodes(css = ".catalog-category-model__price-current-value") %>%
  
  #...and extract the text
  html_text()%>%
  
  # Convert vector to tibble
  enframe(name = "position", value = "price")

#Joining category, model and price
bike_model_tbl <- left_join(bike_model_tbl, model_price_tbl, by = c("position")) %>%
  tibble(category= bike_category_tbl$category_url[2]) %>% 
  separate(col    = category,
           into   = c("remove", "category"),
           sep    = "fahrräder/") %>% 
  select(position, category, model, price)

#Print first 10 rows
bike_model_tbl
```

When you knit this R Markdown document, you will see that the histogram is printed to the page, along with the R code. This document can be set up to hide the R code in the webpage, just delete the comment (hashtag) from the cold folding option in the yaml header up top. For purposes of letting yourself see the code, and me see the code, best to keep it the way that it is. You'll learn that all of these things and more can be customized in each R code block.